Oregon_Wildfire_reduced

### 2017-02-20 r_scripts/oregon_clean.rmd
Update oregon_clean.rmd

---
title: "Oregon_Wildfire_reduced"
author: "Jingyang Liu"
date: "February 15, 2017"
output: html_document
---

@@

```{r}
library(dplyr)
library(data.table)
getwd()

# import data ------------------------------------------------------------------
# Read in first 2000000 lines for exploration.

@@

# I wrote a reduced dataframe of 2000,000 rows to get an idea of data structure

# import the reduced dataframe
start_time <- Sys.time()
oregon_df <- fread("../data/oregon_epis_care_reduced.csv", sep = ",")
stop_time <-  Sys.time() - start_time

@@

## The overall summary  of data set
72 variables in total.
```{r overall summary}

# Explore data structure
summary(oregon_df)
```

## 1) *personkey*, Unique person identifier
person's ID, maximum is 14381640
numeric; length: 8
```{r personkey}
summary(as.factor(oregon_df$personkey))
min(oregon_df$personkey)
max(oregon_df$personkey) # seems length is 8
```

>>>>

---
title: "Oregon_Wildfire_reduced"
author: "Jingyang Liu"
date: "February 15, 2017"
---

@@

```{r}
library(dplyr)
library(data.table)
library(knitr)
getwd()

# import data ------------------------------------------------------------------
# Read in first 200,000 lines for exploration.

@@

# I wrote a reduced dataframe of 200,000 rows to get an idea of data structure

# import the reduced dataframe
start_time <- Sys.time()
oregon_df <- fread("../data/oregon_epis_care_reduced.csv", sep = ",", 
                   colClasses = rep("character", 72))
stop_time <-  Sys.time() - start_time

@@

## The overall summary  of data set
72 variables in total.
```{r overall summary}
# Check data structure
# summary(oregon_df)
```

## convert variable type to appropriate functional form 
(i.e. zipcode leave as character, age as numeric, date as date, dx leave as character, but *NULL* == NA)
``` {r convert variable}
oregon_df$yob <- as.numeric(oregon_df$yob)

oregon_df[oregon_df=="*NULL*"]<- NA
```

# summary statistics

## 1. personkey
```{r}
personkey2 <- oregon_df %>% 
  select(personkey, gender) %>%
  filter(gender=="U") %>%
  mutate(id_ind =ifelse(is.na(personkey)==0 , 1, 
                        ifelse(is.na(personkey)==1, 0, NA))
   )

# table ----
knitr::kable(personkey2)

# preserve order of group
personkey2$group <- reorder(personkey2$id_ind, personkey2$personkey)

## Age
```{r}
oregon_df$yob <- as.numeric(oregon_df$yob)
age2 <- oregon_df %>%
  select(yob) %>%
  mutate(age = 2013-yob) %>%
  mutate(age_ind = ifelse(age < 15, 0,
                   ifelse(age >= 15 & age < 65, 1,
                   ifelse(age >= 65 & age <=120, 2, NA)))
  ) %>%
  group_by(age_ind) %>% 
  summarise(n = n()) %>% 
  mutate(total_n = sum(n),
         percent = round(n/total_n,3)*100,
         age_cat = ifelse(age_ind == 0, "Age <15",
                   ifelse(age_ind == 1, "Age 15 to 65",
                   ifelse(age_ind == 2, "Age >= 65",
                   ifelse(is.na(age_ind), "Missing", NA))))) %>% 
  select(age_cat, age_ind, n, percent)

age2[4,1] <- age2[4,2] <- "Missing"

knitr::kable(age2, caption = "Age")

## summary statistics key variables for study 
(date, dx, need to know urgent and emergency department visits, sex, age, patientid, state of residence, zipcode, county, race)
```{r}
summary(as.factor(oregon_df$personkey))
summary(as.factor(oregon_df$patid))
summary(as.factor(oregon_df$gender))
summary(as.factor(2013-oregon_df$yob)) # birth year -> age
summary(as.factor(oregon_df$race))
summary(as.factor(oregon_df$ethn))
summary(as.factor(oregon_df$STATE))
summary(as.factor(oregon_df$ZIP))
summary(as.factor(oregon_df$MSA)) 
summary(as.factor(oregon_df$fromdate))
summary(as.factor(oregon_df$todate))
summary(as.factor(oregon_df$pos)) # 23
summary(as.factor(oregon_df$megtype)) # Acute for emergency?
summary(as.factor(oregon_df$los))

## 1) *personkey*, Unique person identifier
person's ID, maximum is 14381640
length: 8
```{r personkey}
summary(as.factor(oregon_df$personkey))
min(oregon_df$personkey)
max(oregon_df$personkey) # seems length is 8
```
